---
title: "Project"
author: "Olufunke YUSUF"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
# Library for modeling
library(tidymodels)

# Load tidyverse
library(tidyverse)
```

```{r}
URL <- 'https://dax-cdn.cdn.appdomain.cloud/dax-noaa-weather-data-jfk-airport/1.1.4/noaa-weather-sample-data.tar.gz'

# Download the file
download.file(URL, destfile = "noaa-weather-sample-data.tar.gz")

# Untar the downloaded file
untar("noaa-weather-sample-data.tar.gz", exdir = ".")
```


```{r}
# Read the CSV file
noaa_weather <- read_csv("C:/Users/HP/Downloads/noaa-weather-sample-data (1)/noaa-weather-sample-data/jfk_weather_sample.csv")

# Display the first few rows of the data
head(noaa_weather)
```


```{r}
glimpse(noaa_weather)
```


```{r}
tail(noaa_weather)
```


```{r}
dim(noaa_weather)
```


```{r}
colnames(noaa_weather)
```

Preprocessing Data
```{r}
noaa_weather %>%
  select(HOURLYRelativeHumidity,HOURLYDRYBULBTEMPF,HOURLYPrecip,HOURLYWindSpeed,HOURLYStationPressure)
```


```{r}
unique(noaa_weather$HOURLYPrecip)
```

```{r}
noaa_weather %>%
  HP<-select(HOURLYPrecip)
```


```{r}
# Remove "s" from HOURLYPrecip column
noaa_weather <- noaa_weather %>%
  mutate(HOURLYPrecip = str_remove(HOURLYPrecip, "s"))
```


```{r}
noaa_weather %>%
  select(HOURLYPrecip)
```


```{r}
# Replace "T" with "0.00" in HOURLYPrecip column
noaa_weather <- noaa_weather %>%
  mutate(HOURLYPrecip = ifelse(HOURLYPrecip == "T", "0.00", HOURLYPrecip))
```


```{r}
noaa_weather%>%
  select(HOURLYPrecip)
```


```{r}
glimpse(noaa_weather)
```


```{r}
# Convert HOURLYPrecip to numeric type
noaa_weather <- noaa_weather %>%
  mutate(HOURLYPrecip = as.numeric(HOURLYPrecip))
```

```{r}
head(noaa_weather)
```


```{r}
# Rename specified columns
noaa_weather <- noaa_weather %>%
  rename(
    relative_humidity = HOURLYRelativeHumidity,
    dry_bulb_temp_f = HOURLYDRYBULBTEMPF,
    precip = HOURLYPrecip,
    wind_speed = HOURLYWindSpeed,
    station_pressure = HOURLYStationPressure
  )

# Display the modified data with renamed columns
cat("\nModified data with columns renamed:\n")
head(noaa_weather)
```


```{r}
set.seed(1234)
noaa_split <- initial_split(noaa_weather)
train_data <- training(noaa_split)
test_data <- testing(noaa_split)
```


```{r}
noaa_split2 <- initial_split(noaa_weather, prop = 0.8)  # prop = 0.8 works as well
train_data2 <- training(noaa_split2)
test_data2 <- testing(noaa_split2)
```


```{r}
# Plot histograms for the specified variables
train_data %>%
  ggplot(aes(x = relative_humidity)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Relative Humidity", x = "Relative Humidity", y = "Frequency")
```


```{r}
train_data %>%
  ggplot(aes(x = dry_bulb_temp_f)) +
  geom_histogram(binwidth = 2, fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Dry Bulb Temperature (F)", x = "Dry Bulb Temperature (F)", y = "Frequency")
```


```{r}
train_data %>%
  ggplot(aes(x = precip)) +
  geom_histogram(binwidth = 0.1, fill = "coral", color = "black") +
  labs(title = "Histogram of Precipitation", x = "Precipitation", y = "Frequency")
```


```{r}
train_data %>%
  ggplot(aes(x = wind_speed)) +
  geom_histogram(binwidth = 2, fill = "lightcoral", color = "black") +
  labs(title = "Histogram of Wind Speed", x = "Wind Speed", y = "Frequency")
```


```{r}
train_data %>%
  ggplot(aes(x = station_pressure)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Station Pressure", x = "Station Pressure", y = "Frequency")
```


```{r}
# List of predictor variables
predictors <- c("relative_humidity", "dry_bulb_temp_f", "wind_speed", "station_pressure")
```

```{r}
# Pick linear regression
lm_spec <- linear_reg() %>%
  # Set engine
  set_engine(engine = "lm")

# Print the linear function
lm_spec
```
```{r}
train_fit <- lm_spec %>% 
    fit(precip~relative_humidity, data = train_data)

train_fit
```
```{r}
```{r}
# Visualize the model with a scatter plot
  ggplot(train_data, aes_string(x = relative_humidity, y = "precip")) +
    geom_point(color = "blue") +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(title = paste("Simple Linear Regression: precip ~", relative_humidity),
         x = "Relative Humidity",
         y = "Precipitation")
```

```{r}
train_fit <- lm_spec %>% 
    fit(precip~dry_bulb_temp_f, data = train_data)

train_fit
```


```{r}
train_fit <- lm_spec %>% 
    fit(precip~wind_speed, data = train_data)

train_fit
```
```{r}
train_fit <- lm_spec %>% 
    fit(precip~station_pressure, data = train_data)

train_fit
```


```{r}
predictors <- c("relative_humidity", "dry_bulb_temp_f", "wind_speed", "station_pressure")

options(repr.plot.width = 10, repr.plot.height = 8) # Adjust the plot size if needed
par(mfrow = c(1, 2))

for (predictor in predictors) {
  formula <- formula(paste("precip ~", predictor))
  lm_model <- lm(formula, data = train_data)
  plot(train_data[[predictor]], train_data$precip, main = paste("Scatter Plot: precip ~", predictor),
       xlab = predictor, ylab = "Precipitation", col = "blue")
  abline(lm_model, col = "red")
  
    cat("\nLinear Regression Model Summary for", predictor, ":\n")
  print(summary(lm_model))
}

par(mfrow = c(1, 1))

```


```{r}
# Print the coefficients of the linear regression model
  cat("\nLinear Regression Model Summary for", predictor, ":\n")
  print(summary(lm_model))

```
```{r}
# Load necessary libraries
library(tidyverse)
library(tidymodels)

# Model 1: Add more features
model1_formula <- formula("precip ~ relative_humidity + dry_bulb_temp_f + wind_speed + station_pressure")
lm_model1 <- lm(model1_formula, data = train_data)
```


```{r}
# Model 2: Add regularization using tidymodels (L1 regularization with glmnet)
set.seed(1234)
model2_spec <- linear_reg() %>%
  set_engine("glmnet") %>%
  set_mode("regression")

model2_recipe <- recipe(precip ~ relative_humidity + dry_bulb_temp_f + wind_speed + station_pressure, data = train_data)
```


```{r}
results %>% collect_metrics()
```


```{r}
# Model 1: Linear Regression with More Features
test_predictions1 <- predict(lm_model1, newdata = test_data)
rmse_test <- sqrt(mean((test_predictions1 - test_data$precip)^2))
r_squared1_test <- cor(test_predictions1, test_data$precip)^2
```


```{r}
ggplot(cars, aes(x = relative_humidity, y = precipe)) + 
    geom_point() + 
    geom_smooth(method = "lm", 
                formula = y ~ poly(x, 8), 
                col = "red", se = FALSE) 
```


```{r}
# Model 3: (Add more models as needed)

# Create a comparison table
model_names <- c("Model 1 (More Features)", "Model 2 (Regularized Linear Regression)")
train_rmse <- c(rmse1, rmse2)
test_rmse <- c(rmse1_test, rmse2_test)
train_r_squared <- c(r_squared1, r_squared2)
test_r_squared <- c(r_squared1_test, r_squared2_test)

comparison_df <- data.frame(model_names, train_rmse, test_rmse, train_r_squared, test_r_squared)
```

```{r}

```

```{r}

```

